Hybrid Threads for the Parrot Virtual Machine

### 2012-06-05 Paper.tex
Fix little style problem
Abstracts

\begin{german}
\chapter{Kurzfassung (1)}
\end{german}

@@

moretasks.pir is an extended test where the main task creates 50000 child tasks. These tasks would poll a shared variable with pauses of 100ms. When the variable is set to 1, they schedule tasks on the main thread to write their number to a results array. When the results array contains the expected number of results the process is repeated. This test stresses reading and writing to shared data, allocation and garbage collection. This test proved to be valuable for finding all sorts of concurrency issues. The complete source code is available in appendix \ref{app:moretasks}.

>>>>

Parrot is a runtime for dynamically typed programming languages. Despite several attempts in it's 10 years of history, it does provide any support for multithreaded computation.

This paper shows a way to implement threading support in the runtime using a hybrid between lightweight (``green'') threads and operating system threads. These lightweight threads are used as messages in a system where reading shared data is allowed but only the one owner thread may write to it. The implementation is described in detail and benchmarks presented demonstrating the viability of this design.

\begin{german}
\chapter{Kurzfassung (1)}

Parrot ist eine Laufzeitumgebung f{\"u}r dynamisch typisierte Programmiersprachen. Trotz mehrerer Anl{\"a}ufe in seiner zehnj{\"a}hrigen Geschichte, bietet es keine Unterst{\"u}tzung f{\"u}r Multithreading.

Diese Arbeit zeigt einen Weg um Threadingunterst{\"u}tzung in der Laufzeitumgebung zu implementieren mit Hilfe eines hybriden Systems aus leichtgewichtigen und Betriebssystem- Threads. Diese leichtgewichtigen Threads werden als Nachrichten in einem System benutzt, in dem das Lesen von gemeinsamen Daten erlaubt ist, aber nur der eine besitzende Thread darauf schreiben darf. Die Implementierung wird detailiert beschrieben und mit Benchmarks die Brauchbarkeit des Designs gezeigt.

\end{german}

@@

moretasks.pir is an extended test where the main task creates 50000 child tasks. These tasks would poll a shared variable with pauses of 100ms. When the variable is set to 1, they schedule tasks on the main thread to write their number to a results array. When the results array contains the expected number of results the process is repeated.

This test stresses reading and writing to shared data, allocation and garbage collection. It proved to be valuable for finding all sorts of concurrency issues. The complete source code is available in appendix \ref{app:moretasks}.
